An Efficient Forecasting Approach for the Real-Time Reduction of Boundary Effects\\ in Time-Frequency Representations

### 2020-07-22 Notes/BoundEffRed.tex
result

\usepackage{amsmath,amssymb,amsfonts,amsthm}
\usepackage{euscript}
\usepackage[latin1]{inputenc}
\usepackage[T1]{fontenc}
\usepackage[english]{babel}
\usepackage{color}
\usepackage{graphicx}
\usepackage{mathrsfs}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{dsfont}
\usepackage[shortlabels]{enumitem}
\usepackage{multirow}
\usepackage{hhline}
\usepackage{hyperref}
\usepackage{mathpazo}
\usepackage{array}
\usepackage{bigints}

@@

\section{Conclusion}
\label{se:conclusion}
In this paper, we propose an algorithm, named {\sf BoundEffRed}, for the real-time reduction of boundary effects in TF representations. This method is based on an extension of the signal obtained by a simple-minded and numerically efficient forecasting. We have shown theoretically that the chosen dynamic model is sufficient to extend signals formed by a sum of sine waves. Moreover, thanks to the low computational time, the numerical results show that this strategy allows us to switch to a real-time implementation of {\sf BoundEffRed}, unlike other existing forecasting methods. The numerical results also confirmed the robustness to noise of {\sf BoundEffRed}, as well as its ability to be applied to many time-frequency representations. 
%Other numerical examples are also included, such as the application of {\sf BoundEffRed} to other types of biomedical signals such as ECGs.

>>>>

\usepackage{amsmath,amssymb,amsfonts,amsthm}
\usepackage{euscript}
\usepackage[latin1]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{color}
\usepackage{graphicx}
\usepackage{mathrsfs}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{dsfont}
\usepackage[shortlabels]{enumitem}
\usepackage{multirow}
\usepackage{hhline}
\usepackage{hyperref}
\usepackage{mathpazo}
\usepackage{array}
\usepackage{bigints}

@@

\section{Conclusion}
\label{se:conclusion}
In this paper, we propose an algorithm, named {\sf BoundEffRed}, for the real-time reduction of boundary effects in TF representations. This method is based on an extension of the signal obtained by a simple-minded and numerically efficient forecasting. We have shown theoretically that the chosen dynamic model is sufficient to extend signals formed by a sum of sine waves. Moreover, thanks to the low computational time, the numerical results show that this strategy allows us to switch to a real-time implementation of {\sf BoundEffRed}, unlike other existing forecasting methods. The numerical results also confirmed the robustness to noise of {\sf BoundEffRed}, as well as its ability to be applied to many TF representations. 
%Other numerical examples are also included, such as the application of {\sf BoundEffRed} to other types of biomedical signals such as ECGs.

### 2020-07-22 Notes/results.tex
new XP
result

In the aim of comparing the different techniques, we use a criterion, proposed in~\cite{Daubechies16conceft}, that quantifies the distance between a given TF representation and the optimal one. It is built in analogy with the optimal transport distance, which enables quantifying the distance between two probability density functions. Let us generically denote a time-frequency representation $\ccQ$. Then, for $t$ fixed, we consider the following probability density function: $p_\ccQ^t(\xi) = \left. |\ccQ(\xi,t)|^2 \middle/ \int_\RR |\ccQ(\nu,t)|^2\dd\nu \right.$. At each instant $t$, we then determine the optimal transport distance $d_{t}$ between the two densities. It is given by the $L^1$ norm of the difference between the associated distribution functions. In other words, we have:
\begin{equation*}
d_{t}(\ccQ,\ccF_0) = \int_\RR\left|\tilde P_{\ccQ}^t(\xi)-  P_{\ccF_0}^t(\xi)\right|\dd\xi\ ,
\end{equation*}
where $P_{\ccQ}^t(\xi)=\int_{-\infty}^\xi p_{\ccQ}^t(\nu)\dd\nu$ and $\tilde P_{\ccF_0}^t(\xi)=\int_{-\infty}^\xi\tilde p_{\ccF_0}^t(\nu)\dd\nu$. The \textit{Optimal Transport Distance (OTD)} quantifies the proximity between the estimated and actual instantaneous frequencies while favoring the sparsity of the estimated TF representation. That is why, the performance index $D(\ccQ)$ for the reduction of boundary effects of a given TF representation $\ccQ$ is given by the ratio between its averaged OTD to the optimal TF representation $\ccF^\mathrm{opt}$ (see~\eqref{eq:opt.TFR}) and the averaged OTD of the original TF representation $\ccF$ to $\ccF^\mathrm{opt}$, that is:
%\begin{equation}
%D(\ccQ) = \dfrac1{|I|}\bigintssss_I\ \frac{d_{t}\left(\ccQ,\ccF^\mathrm{opt}\right)}{d_{t}\left(\ccF,\ccF^\mathrm{opt}\right)}\ \dd t\ .
%\label{eq:index.perf}
%\end{equation}
\begin{equation}
D(\ccQ) = \dfrac{\displaystyle\int_I d_{t}\left(\ccQ,\ccF^\mathrm{opt}\right)\dd t}{\displaystyle\int_I d_{t}\left(\ccF,\ccF^\mathrm{opt}\right)\dd t}\ .
\label{eq:index.perf}
\end{equation}
Thus, $D(\ccQ)<1$ means a reduction of the boundary effects. Let us evaluate the quality of the boundary effects reduction on biomedical signals.

@@

\subsubsection{Photoplethysmogram}
\label{ssse:ppg}
We perform a study similar to the previous one on a 640 second-long photoplethysmogram (PPG) signal extracted from the Physionet dataset~\cite{Pimentel17toward, Goldberger00physiobank}, sampled at $\fs=125$~Hz. A portion of this signal is displayed in Fig.~\ref{fig:ppg}. The estimated 5 seconds extensions of this segment obtained by {\sf SigExt}, EDMD, and GPR forecastings  are superimposed to the ground-truth extension in Fig.~\ref{fig:ppg}.

@@

\begin{table}
\centering
\caption{PPG Signal: Performance of the Boundary-Free TF Representations According to the Extension Method}
\begin{tabular}{|c||c|c|c|}
  \hline
   \multirow{2}{40pt}{\centering Extension method} & \multicolumn{3}{c|}{Averaged performance index $D$} \\
   \cline{2-4}
      & STFT & SST & ConceFT\\
   \hhline{|=#=|=|=|}
   {\sf SigExt} & $0.280$ & $0.309$ & $0.367$ \\
   \hline
   Symmetric & $1.168$ & $1.209$ & $1.310$ \\
   \hline
   EDMD & $0.289$ & $0.319$ & $0.375$ \\
   \hline
   GPR & $0.276$ & $0.303$ & $0.361$ \\
   \hline
\end{tabular}
\label{tab:otd.ppg}
\end{table}

@@

\begin{figure}
\centering
\includegraphics[width=.48\textwidth]{OTDBoundEffRed.eps}
\caption{PPG signal. Averaged OTD of {\sf BoundEffRed} in function of the SNR.}
\label{fig:otd.noise}
\end{figure}

>>>>

In the aim of comparing the different techniques, we use a criterion, proposed in~\cite{Daubechies16conceft}, that quantifies the distance between a given TF representation and the optimal one. It is built in analogy with the optimal transport distance, which enables quantifying the distance between two probability density functions. Let us generically denote a TF representation $\ccQ$. Then, for $t$ fixed, we consider the following probability density function: $p_\ccQ^t(\xi) = \left. |\ccQ(\xi,t)|^2 \middle/ \int_\RR |\ccQ(\nu,t)|^2\dd\nu \right.$. At each instant $t$, we then determine the optimal transport distance $d_{t}$ between the two densities. It is given by the $L^1$ norm of the difference between the associated distribution functions. In other words, we have:
\begin{equation*}
d_{t}(\ccQ,\ccF_0) = \int_\RR\left|\tilde P_{\ccQ}^t(\xi)-  P_{\ccF_0}^t(\xi)\right|\dd\xi\ ,
\end{equation*}
where $P_{\ccQ}^t(\xi)=\int_{-\infty}^\xi p_{\ccQ}^t(\nu)\dd\nu$ and $\tilde P_{\ccF_0}^t(\xi)=\int_{-\infty}^\xi\tilde p_{\ccF_0}^t(\nu)\dd\nu$. The \textit{Optimal Transport Distance (OTD)} quantifies the proximity between the estimated and actual instantaneous frequencies while favoring the sparsity of the estimated TF representation. That is why, the performance index $D(\ccQ)$ for the reduction of boundary effects of a given TF representation $\ccQ$ is given by the ratio between its averaged OTD to the optimal TF representation $\ccF^\mathrm{opt}$ (see~\eqref{eq:opt.TFR}) and the averaged OTD of the original TF representation $\ccF$ to $\ccF^\mathrm{opt}$, that is:
%\begin{equation}
%D(\ccQ) = \dfrac1{|I|}\bigintssss_I\ \frac{d_{t}\left(\ccQ,\ccF^\mathrm{opt}\right)}{d_{t}\left(\ccF,\ccF^\mathrm{opt}\right)}\ \dd t\ .
%\label{eq:index.perf}
%\end{equation}
\begin{equation}
D(\ccQ) = \dfrac{\displaystyle\int_I d_{t}\left(\ccQ,\ccF^\mathrm{opt}\right)\dd t}{\displaystyle\int_I d_{t}\left(\ccF,\ccF^\mathrm{opt}\right)\dd t}\ .
\label{eq:index.perf}
\end{equation}
Thus, $D(\ccQ)<1$ means a reduction of the boundary effects. Let us evaluate the quality of the boundary effects reduction on biomedical signals.

@@

\subsubsection{Photoplethysmogram}
\label{ssse:ppg}
We perform a study similar to the previous one on a $640$-second-long photoplethysmogram (PPG) signal extracted from the Physionet dataset~\cite{Pimentel17toward, Goldberger00physiobank}, sampled at $\fs=125$~Hz. A portion of this signal is displayed in Fig.~\ref{fig:ppg}. The estimated 5 seconds extensions of this segment obtained by {\sf SigExt}, EDMD, and GPR forecastings  are superimposed to the ground-truth extension in Fig.~\ref{fig:ppg}.

@@

\begin{table}
\centering
\caption{PPG Signal: Performance of the Boundary-Free TF Representations According to the Extension Method}
\begin{tabular}{|c||c|c|c|c|}
  \hline
   \multirow{2}{40pt}{\centering Extension method} & \multicolumn{4}{c|}{Averaged performance index $D$} \\
   \cline{2-5}
      & STFT & SST & ConceFT & RS\\
   \hhline{|=#=|=|=|=|}
   {\sf SigExt} & $0.280$ & $0.309$ & $0.367$ & $0.534$ \\
   \hline
   Symmetric & $1.168$ & $1.209$ & $1.310$ & $0.983$ \\
   \hline
   EDMD & $0.289$ & $0.319$ & $0.375$ & $0.503$ \\
   \hline
   GPR & $0.276$ & $0.303$ & $0.361$ & $0.544$ \\
   \hline
\end{tabular}
\label{tab:otd.ppg}
\end{table}

@@

\begin{figure}
\centering
\includegraphics[width=.48\textwidth]{OTDBoundEffRed.eps}
\caption{Performance index of {\sf BoundEffRed} applied on a PPG, in function of the SNR.}
\label{fig:otd.noise}
\end{figure}
